New lower bounds for Schur and weak Schur numbers

### 2021-06-06 article/schur.tex
corrections partie schur

@@

\begin{theorem}
	Let \((p,k), (q,n) \in (\mathbb{N}^*)^2\). If there exists a \(S\)-template with width \(q\) and \(n+1\) colors 
	and a partition of \([\![1,p]\!]\) into \(k\) sum-free subsets then there exists a partition of \([\![1,pq+m_{n+1}
	-1]\!]\) into \(n+k\) sum-free subsets. \(m_{n+1}\) denotes the minimum element colored with the special color in the \(S\)-template.
\end{theorem}

@@

\begin{corollary}
	Let \(n, k \in \mathbb{N}^*\). Then
	\[ S(n+k) \geqslant S^+(n+1)S(k) + m_{n+1} - 1 \]
\end{corollary}

@@

We now proceed to prove the above mentioned theorem.

@@

The following proposition may improve the additive constant of a S-template.

@@

Then, for every \(n \in \mathbb{N}^*\), by using on the last row the coloring \(x \longmapsto g(x - p S(n))\), we have
\[ S(n+k) \geqslant S^+(n+1)S(k) + b\].
\end{proposition}

This proposition corresponds to the fact that sometimes a column is not the sum of two columns of a given color, but 
adding this column to the color would create sums in the color when applying the extension procedure. However, the 
last row does not intereact with the right-most columns when it comes to creating new sums. As a result, the hypotheses 
made on the coloring of the last row can be weakened.

@@

\begin{proof}[\textsc{Proof.}]
The idea is the same as in the previous theorem. The only difference is the S-template property inherited 
from the second S-template. \\
\end{proof}

\subsection{Inequalities and new lower bounds for Schur numbers}

@@

\par
Here are the best inequalities on Schur numbers so far (the \hyperref[S-templates]{templates} corresponding to 
the third, fourth and fifth inequalities can be found in the appendix).
\[
\begin{array}{r r r r r}
	S(n + 1) & \geqslant & 3 S(n) & + & 1 \\
	S(n + 2) & \geqslant & 9 S(n) & + & 4 \\
	S(n + 3) & \geqslant & 33 S(n) & + & 6 \\
	S(n + 4) & \geqslant & 111 S(n) & + & 43 \\
	S(n + 5) & \geqslant & 380 S(n) & + & 148 \\
	S(n + 6) & \geqslant & 1\,140 S(n) & + & 528 \\
\end{array}
\]

The first inequality comes from  Schur's original article\cite{Schur1917}. The second one is due to
Abott \cite{AbbottHanson}
and the third one to Rowley \cite{RowleyRamsey}. The three last inequalies are our result.

\par
The first three inequalities are optimal. The fourth one is optimal among symmetric S-templates whose special color is 
the last in the order of apparition (and with a multiplicative factor less than or equal to 118). The fifth one is most 
likely not optimal but should not be too far from the optimum. 
Finally, the sixth one is obtained by combining a S-template with width 380 and one with width 3. 
Although we could not find a better S-template with seven colors, the last inequality is definitely very far from the 
optimal value. One may try to seek better templates by constraining less the search space and by using 
Monte-Carlo methods, as in \cite{Bouzy2015AnAP}. This could be the subject of a future work.

@@

\[
\begin{array}{c}
	\resetarraystretch
	\begin{NiceArray}{cwc{8ex}wc{10ex}wc{10ex}wc{11ex}}[hvlines]
	\CodeBefore
		\cellcolor{yellow}{2-2}
		\cellcolor{yellow}{3-3}
		\cellcolor{yellow}{4-4}
		\cellcolor{yellow}{4-5}
	\Body
		n & 8 & 9 & 10 & 11 \\
		33 \, S(n-3) + 6 & 5\,286 & 17\,694 & 55\,446 & 174\,444 \\
		111 \, S(n-4) + 43 & 4927 & 17\,803 & 59\,539 & 186\,523 \\
		380 \, S(n-5) + 148 & 5\,088 & 16\,868 & 60\,948 & 203\,828 \\
		1\,140 \, S(n-6) + 528 & 5\,088 & 15\,348 & 50\,688 & 182\,928 \\
	\end{NiceArray}
	\\ \\
	\resetarraystretch
	\begin{NiceArray}{cwc{8ex}wc{10ex}wc{10ex}wc{11ex}}[hvlines]
	\CodeBefore
		\cellcolor{yellow}{2-3}
		\cellcolor{yellow}{4-2}
		\cellcolor{yellow}{4-4}
		\cellcolor{yellow}{4-5}
	\Body
		n & 12 & 13 & 14 & 15 \\
		33 \, S(n-3) + 6 & 587\,505 & 2\,011\,290 & 6\,726\,330 & 21\,072\,090 \\
		111 \, S(n-4) + 43 & 586\,789 & 1\,976\,176 & 6\,765\,271 & 22\,624\,951 \\
		380 \, S(n-5) + 148 & 638\,548 & 2\,008\,828 & 6\,765\,288 & 23\,160\,388 \\
		1\,140 \, S(n-6) + 528 & 611\,568 & 1\,915\,728 & 6\,026\,568 & 20\,295\,948 \\
	\end{NiceArray}
\end{array}
\]

>>>>

We define \(S^+\) to reduce the constraints of Abbott and Hanson's construction to improve this construction.

@@

\begin{theorem}
\label{Theorem2.4}
	Let \((p,k), (q,n) \in (\mathbb{N}^*)^2\). If there exists a \(S\)-template with width \(q\) and \(n+1\) colors 
	and a partition of \([\![1,p]\!]\) into \(k\) sum-free subsets then there exists a partition of \([\![1,pq+m_{n+1}
	-1]\!]\) into \(n+k\) sum-free subsets. \(m_{n+1}\) denotes the minimum element colored with the special color in the \(S\)-template.
\end{theorem}

@@

\begin{corollary}
\label{Corollary2.5}
	Let \(n, k \in \mathbb{N}^*\). Then
	\[ S(n+k) \geqslant S^+(n+1)S(k) + m_{n+1} - 1 \]
\end{corollary}

@@

We now proceed to prove \hyperref[Theorem2.4]{Theorem 2.4}

@@

The following proposition may improve the additive constant of \hyperref[Corollary2.5]{Corollary 2.5}

@@

Then, for every \(n \in \mathbb{N}^*\), by using on the last row the coloring \(x \longmapsto g(x - p S(n))\), we have
\[ S(n+k) \geqslant S^+(n+1)S(k) + b\].
\end{proposition}
This proposition corresponds to the fact that the hypotheses made on the coloring of the last row can be weakened and as a result, we can change the coloring on the last row to extend the previous partitions.
\newline

@@

\begin{proof}[\textsc{Proof.}]
The idea is the same as in the \hyperref[Theorem2.4]{Theorem 2.4}. The only difference is the S-template property inherited 
from the second S-template. \\
\end{proof}

\subsection{New lower bounds for Schur numbers}
In the following, we give the strongest inequalities we have so far and further insight about how they were produced.

@@

\par
Here are the best inequalities on Schur numbers so far (the \hyperref[S-templates]{templates} corresponding to 
the third, fourth and fifth inequalities can be found in the appendix).

\begin{align}
	S(n + 1) \geqslant  3 S(n)  + 1 \label{S(n+1)}\\
	S(n + 2) \geqslant  9 S(n)  + 4 \label{S(n+2)}\\
	S(n + 3) \geqslant  33 S(n) + 6 \label{S(n+3)}\\
	S(n + 4) \geqslant  111 S(n) + 43 \label{S(n+4)}\\
	S(n + 5) \geqslant  380 S(n) + 148 \label{S(n+5)}\\
	S(n + 6) \geqslant 1\,140 S(n) + 528 \label{S(n+6)}
\end{align}

\hyperref[S(n+1)]{(\ref{S(n+1)})} comes from  Schur's original article\cite{Schur1917}. \hyperref[S(n+2)]{(\ref{S(n+2)})} is due to
Abott \cite{AbbottHanson}
and \hyperref[S(n+3)]{(\ref{S(n+3)})} to Rowley \cite{RowleyRamsey}. \hyperref[S(n+4)]{(\ref{S(n+4)})}, \hyperref[S(n+5)]{(\ref{S(n+5)})} and \hyperref[S(n+6)]{(\ref{S(n+6)})} are our result.

\par
\hyperref[S(n+1)]{(\ref{S(n+1)})}, \hyperref[S(n+2)]{(\ref{S(n+2)})} and \hyperref[S(n+3)]{(\ref{S(n+3)})} are optimal. \hyperref[S(n+4)]{(\ref{S(n+4)})} is optimal among symmetric S-templates whose special color is 
the last in the order of apparition (and with a multiplicative factor less than or equal to 118). \hyperref[S(n+5)]{(\ref{S(n+5)})} is most 
likely not optimal but should not be too far from the optimum. 
Finally, \hyperref[S(n+6)]{(\ref{S(n+6)})} is obtained by combining a S-template with width 380 and one with width three. 
Although we could not find a better S-template with seven colors, \hyperref[S(n+6)]{(\ref{S(n+6)})} is definitely very far from the 
optimal value. One may try to seek better templates by constraining less the search space and by using 
Monte-Carlo methods, as in \cite{Bouzy2015AnAP}. This could be the subject of a future work.

@@

\begin{table}[H]
\caption{New lower bounds for \( n \in [\![8,15]\!] \)}
\label{LowerBoundsWS}
\[
\begin{array}{c}
	\resetarraystretch
	\begin{NiceArray}{cwc{8ex}wc{10ex}wc{10ex}wc{11ex}}[hvlines]
	\CodeBefore
		\cellcolor{yellow}{2-2}
		\cellcolor{yellow}{3-3}
		\cellcolor{yellow}{4-4}
		\cellcolor{yellow}{4-5}
	\Body
		n & 8 & 9 & 10 & 11 \\
		33 \, S(n-3) + 6 & 5\,286 & 17\,694 & 55\,446 & 174\,444 \\
		111 \, S(n-4) + 43 & 4927 & 17\,803 & 59\,539 & 186\,523 \\
		380 \, S(n-5) + 148 & 5\,088 & 16\,868 & 60\,948 & 203\,828 \\
	\end{NiceArray}
	\\ \\
	\resetarraystretch
	\begin{NiceArray}{cwc{8ex}wc{10ex}wc{10ex}wc{11ex}}[hvlines]
	\CodeBefore
		\cellcolor{yellow}{2-3}
		\cellcolor{yellow}{4-2}
		\cellcolor{yellow}{4-4}
		\cellcolor{yellow}{4-5}
	\Body
		n & 12 & 13 & 14 & 15 \\
		33 \, S(n-3) + 6 & 587\,505 & 2\,011\,290 & 6\,726\,330 & 21\,072\,090 \\
		111 \, S(n-4) + 43 & 586\,789 & 1\,976\,176 & 6\,765\,271 & 22\,624\,951 \\
		380 \, S(n-5) + 148 & 638\,548 & 2\,008\,828 & 6\,765\,288 & 23\,160\,388 \\
	\end{NiceArray}
\end{array}
\]
\end{table}
\resetarraystretch
